---
title: "Data Types"
author: "Akis"
date: "1/3/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<center> <h1>Objects and Attributes</h1> </center>

### Five basic atomic classes of objects

* character (or string)

* Numeric 

* Integer

* Complex

* Logical (True/False)

### Object Attributes

The properties associated with an object is called an attribute. These properties explain what the object represents.

* names, dimnames

* dimensions (e.g., arrays, matrices)

* class

* length

* user-defined attributes/metadata

The attributes of an object can be accessed with the ```attributes()``` function.

### Objects

* <font size="3"> Vector </font>\
A sequence of elements of the **same data type**. Create vectors with the functions ```vector()``` and ```c()```.\

```{r}
x <- vector("numeric", length = 5) 
```
```{r}
x <- c("a" , "b", "c") #character
x <- c(0.7, 0.8) #numeric
x <- 3:9 #integer
x <- c(1+0i, 2+4i) #complex
x <- c(T, F) #logical
```



```c()```is also used to create vectors.

* <font size="3"> List </font> ```<-``` A sequence of elements of **different data types**.\
Corresponding function is ```list()```.

* <font size="3"> Matrix </font> ```<-``` A collection of elements of the **same data type arranged into a fixed number of rows and columns**.\
Corresponding function is ```matrix()```.

* <font size="3"> Array </font> ```<-``` An array can have **one, two or more dimensions** (see Object Attributes).\
Corresponding function is ```array()```.

* <font size="3"> Factor </font> ```<-``` Used to represent **categorical data** (e.g., males, females).\
Corresponding function is ```factor()```.

* <font size="3"> Data Frame </font> ```<-``` Display data in a **table format** (like SPSS).\
Corresponding function is ```data.frame()```.