---
title: "Data Structures"
author: "Akis"
date: "18/01/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<u><center><h1> Data Structures </u></center></h1>

Any object that contains data is called a data structure.

* <font size="3"> Vector </font>\
A sequence of elements of the **same data type**. Even a single number is considered a vector of length 1. Create vectors with the functions ```vector()``` and ```c()```.\

* <font size="3"> List </font> ```<-``` A sequence of elements of **different data types**.\
Corresponding function is ```list()```.

* <font size="3"> Matrix </font> ```<-``` A collection of elements of the **same data type arranged into a fixed number of rows and columns**.\
Corresponding function is ```matrix()```.

* <font size="3"> Factor </font> ```<-``` Used to represent **categorical data** (e.g., males, females).\
Corresponding function is ```factor()```.

* <font size="3"> Data Frame </font> ```<-``` Display data in a **table format** (like SPSS).\
Corresponding function is ```data.frame()```.

<center><h1> Vectors </center></h1>

The ```c()``` function can be used to create vectors of objects.

```{r}
x <- c(0.5, 0.6)       ## numeric
x <- c(TRUE, FALSE)    ## logical
x <- c(T, F)           ## logical
x <- c("a", "b", "c")  ## character
x <- 9:29              ## integer
x <- c(1+0i, 2+4i)       ## complex
```

Using the ```vector()``` function:
```{r}
x <- vector("numeric", length = 10) #for numeric vectors the    default value is zero.
print(x)
```

Mathematical operations with vectors are performed on each of their elements. For example:
```{r}
my_vector <- c(1, 2, 3)
my_vector + 1 #Adds one to each of the element of my_vector.

my_vector1 <- c(2,4,6)
my_vector + my_vector1 #Adds the 1st element of my_vector to the 1st of my_vector1, the 2nd to the 2nd, etc...
```


### Mixing objects

When different objects are mixed in a vector, *coercion* occurs so that every element in the vector is of the same class.

```{r}
y <- c(1.7, "a")   ## character
y <- c(TRUE, 2)    ## numeric
y <- c("a", TRUE)  ##character
```

### Explicit Coercion

Objects can be explicitly coerced from one class to another using the `as.na*` functions, if available.

```{r}
x <- 0:6
class(x)
as.numeric(x)
as.logical(x)
as.character(x)
```

Non-sensical coercion results in NAs.

```{r}
x <- c("a", "b", "c")
as.numeric(x)
```

<center><h1> Lists </center></h1>

Lists are a special type of vector that can contain elements of different classes. Lists are a very important data type in R.

```{r}
x <- list(1, "a", TRUE, 1+4i)
x
```


<center><h1> Matrices </center></h1>

Matrices are vectors with a dimension attribute. The dimension attribute is itself an integer vector of length 2 (nrow, ncol). Matrices must have every element be the same class.

```{r}
m <- matrix(nrow = 2, ncol = 3)
m
dim(m) #Retrieves the dimensions of m
attributes(m) #Retrieves the attributes of m
```

Matrices can be created directly from vectors by adding a dimension attribute.

```{r}
m <- 1:10
m
dim(m) <- c(2,5) # 2 rows and 5 columns
m
```

### Column-bindings (cbind) and row-binding(rbind)

Matrices can be created by cbind() or rbind().

```{r}
x <- 1:3
y <- 10:13

cbind(x, y)
rbind(x, y)
```

<center><h1> Factors </center></h1>

Factors are used to represent categorical data. Factors can be ordered or un-ordered.

* Factors can be self-explanatory due to labeling.
* Factors are treated specially by modelling functions like **lm()* and *glm()*

```{r}
x <- factor(c("yes", "yes", "no", "yes", "no"))
x
table(x) # Gives frequency count of each level
unclass(x) # This function strips out the class
```

### Factor levels

The order of the levels can be set using the **levels** argument to **factor**. This can be important in linear modelling because the first level is used as the baseline level.

```{r}
x <- factor(c("yes", "no", "yes", "no"),
            levels = c("yes", "no"))
x
```

<center><h1> Data Frames </center></h1>

Data frames are used to store tabular data.

* They are represented as a special type of list where every element of the list has to have the same length.
* Each element of the list can be thought of as a column and the length of each element is the number of rows.
* Unlike matrices, data frames can store different classes of objects in each columns
* Data frames have a special attribute called **row.names**
* Data frames are usually created by calling **read.table()** or **read.csv()**
* Data frames can be converted to a matrix by calling **data.matrix()**

```{r}
x <- data.frame(foo = 1:4, bar = c(T, T, F, F))
x
nrow(x) #returns number of rows
ncol(x) #returns number of columns
```



