---
title: "Data Structures"
author: "Akis"
date: "18/01/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<u><center><h1> Data Structures </u></center></h1>

Any object that contains data is called a data structure.

* <font size="3"> Atomic Vector </font> ```<-``` A sequence of elements of the **same data type**. Even a single number is considered a vector of length 1. Create vectors with the functions ```vector()``` and ```c()```.\

* <font size="3"> List </font> ```<-``` A sequence of elements of **different data types**.\
Corresponding function is ```list()```.

* <font size="3"> Matrix </font> ```<-``` A collection of elements of the **same data type arranged into a fixed number of rows and columns**.\
Corresponding function is ```matrix()```.

* <font size="3"> Factor </font> ```<-``` Used to represent **categorical data** (e.g., males, females).\
Corresponding function is ```factor()```.

* <font size="3"> Data Frame </font> ```<-``` Display data in a **table format** (like SPSS).\
Corresponding function is ```data.frame()```.

<center><h1> Vectors </center></h1>

The ```c()``` function can be used to create vectors of objects.

```{r}
x <- c(0.5, 0.6)       ## numeric
x <- c(TRUE, FALSE)    ## logical vectors can contain the values TRUE, FALSE and NA
x <- c("a", "b", "c")  ## elements of character vectors can be joined together with the paste() function
x <- 9:29              ## integer
x <- c(1+0i, 2+4i)       ## complex
```

Using the ```vector()``` function:
```{r}
x <- vector("numeric", length = 10) #for numeric vectors the default value is zero.
print(x)
```

Mathematical operations with vectors are performed on each of their elements. For example:
```{r}
my_vector <- c(1, 2, 3)
my_vector + 1 #Adds one to each of the element of my_vector.

my_vector1 <- c(2,4,6)
my_vector + my_vector1 #Adds the 1st element of my_vector to the 1st of my_vector1, the 2nd to the 2nd, etc...
```
When given two vectors of the same length, R simply performs the specified arithmetic operation (`+`, `-`, `*`, etc.) element-by-element. If the vectors are of different lengths, R 'recycles' the shorter vector until it is the same length as the longer vector. If the length of the shorter vector does not divide evenly into the length of the longer vector, R will still apply the 'recycling' method, but will throw a warning to let you know something fishy might be going on.

Other relevant functions are:
* `seq()` - creates sequence **from** number x **to** number y, **by** argument z
```{r}
seq(0, 20, by = 10)
```

* `seq_along()` - creates a sequence with the length of another sequence
```{r}
my_seq <- seq(0, 20, by = 10)
my_seq
seq_along(my_seq)
```

* `rep()` - replicates a vector for a number of times, or each element of the vector

```{r}
rep(0, times = 12) #replicates the vector 12 times
rep(c(0, 1, 2), times = 4) #replicated the vector 4 times
rep(c(0, 1, 2), each = 10) #replicates each element of the vector 10 times
```

* `pate()` - joins together the elements of a vector or multiple vectors and converts them to strings
```{r}
example <- c("my", "character", "vector", "is", "this")
paste(example, collapse = " ") #note the space between double quotes

example_cont <- c("but", "I", "don't", "like", "it")
```


### Mixing objects

When different objects are mixed in a vector, *coercion* occurs so that every element in the vector is of the same class.

```{r}
y <- c(1.7, "a")   ## character
y <- c(TRUE, 2)    ## numeric
y <- c("a", TRUE)  ##character
```

### Explicit Coercion

Objects can be explicitly coerced from one class to another using the `as.na*` functions, if available.

```{r}
x <- 0:6
class(x)
as.numeric(x)
as.logical(x)
as.character(x)
```

Non-sensical coercion results in NAs.

```{r}
x <- c("a", "b", "c")
as.numeric(x)
```

<center><h1> Lists </center></h1>

Lists are a special type of vector that can contain elements of different classes. Lists are a very important data type in R.

```{r}
x <- list(1, "a", TRUE, 1+4i)
x
```


<center><h1> Matrices </center></h1>

Matrices are vectors with a dimension attribute. The dimension attribute is itself an integer vector of length 2 (nrow, ncol). Matrices must have every element be the same class.

```{r}
m <- matrix(nrow = 2, ncol = 3)
m
dim(m) #Retrieves the dimensions of m
attributes(m) #Retrieves the attributes of m
```

Matrices can be created directly from vectors by adding a dimension attribute.

```{r}
m <- 1:10
m
dim(m) <- c(2,5) # 2 rows and 5 columns
m
```

### Column-bindings (cbind) and row-binding(rbind)

Matrices can be created by cbind() or rbind().

```{r}
x <- 1:3
y <- 10:13

cbind(x, y)
rbind(x, y)
```

<center><h1> Factors </center></h1>

Factors are used to represent categorical data. Factors can be ordered or un-ordered.

* Factors can be self-explanatory due to labeling.
* Factors are treated specially by modelling functions like **lm()* and *glm()*

```{r}
x <- factor(c("yes", "yes", "no", "yes", "no"))
x
table(x) # Gives frequency count of each level
unclass(x) # This function strips out the class
```

### Factor levels

The order of the levels can be set using the **levels** argument to **factor**. This can be important in linear modelling because the first level is used as the baseline level.

```{r}
x <- factor(c("yes", "no", "yes", "no"),
            levels = c("yes", "no"))
x
```

<center><h1> Data Frames </center></h1>

Data frames are used to store tabular data.

* They are represented as a special type of list where every element of the list has to have the same length.
* Each element of the list can be thought of as a column and the length of each element is the number of rows.
* Unlike matrices, data frames can store different classes of objects in each columns
* Data frames have a special attribute called **row.names**
* Data frames are usually created by calling **read.table()** or **read.csv()**
* Data frames can be converted to a matrix by calling **data.matrix()**

```{r}
x <- data.frame(foo = 1:4, bar = c(T, T, F, F))
x
nrow(x) #returns number of rows
ncol(x) #returns number of columns
```


### Add columns to an existing data frame

```{r}
# my data set
my_matrix <- matrix(1:20, nrow = 4, ncol = 5)
# Column to add
patients <- c("Bill", "Gina", "Kelly", "Sean")

# Add to data set
new_dataset <- cbind(my_matrix, patients)
new_dataset

# Add column names
cnames <- c("pid", "age", "weight", "bp", "rating", "patient")
colnames(new_dataset) <- cnames
new_dataset
```




