---
title: "Data Types and Attributes"
author: "Akis"
date: "1/3/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Five basic atomic classes of objects

* character (or string)

* Numeric 

* Integer

* Complex

* Logical (True/False)

### Object Attributes

The properties associated with an object is called an attribute. These properties explain what the object represents.

* names, dimnames

* dimensions (e.g., arrays, matrices)

* class

* length

* user-defined attributes/metadata

The attributes of an object can be accessed with the ```attributes()``` function.

<center><h1> Names Attribute </center></h1>

R objects can have names, which is very useful for writing readable code and self-describing objects.

```{r}
x <- 1:3
names(x) #returns NULL because we did not give a name to x
names(x) <- c("foo", "bar", "norf")
x
names(x)
```

Lists can also have names.

```{r}
x <- list(a = 1, b = 2, c = 3) # The = operator is used to assign a name
x
```

Matrices can have names too, but they are called dimnames.

```{r}
m <- matrix(1:4, nrow = 2, ncol = 2)
dimnames(m)
dimnames(m) <- list(c("a", "b"), c("c", "d"))
m
```

When using the **dimnames** function, we assign it a list where the first element of the list is a vector of row names and the second element is a vector of column names. So we have named the rows a and b, and named the columns c and d. 

# Give names to pre-existing objects
```{r}
# Unnamed object
vect2 <- c(11, 2, NA)
# Naming the object
names(vect2) <- c("foo", "bar", "norf")
```

