---
title: "Subsetting basics"
author: "Akis"
date: "17/01/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Subsetting objects in R

A number of operators can be used to extract subsets of R objects:

* [] - returns an object of the same class as the original (can be used to select more than one element with one exception...)

For example if you subset a vector you will get back a vector. Subset a list and you will get a list.

```{r, collapse=TRUE}
#Extract the first element of object x
x <- c("a", "b", "c", "d", "e")
x[1]

#Extract the first 3 elements of object x
x[1:3]
```



* [[]] - extracts elements of a list or data frame (can only be used to extract a single element but the returned object will not necessarily be a list or data frame)

Remember lists can hold objects of different classes.

* $ - extracts elements of a list or data frame by name

# Subsetting vectors
How to extract elements from a vector based on specified conditions:

For example, we may only be interested in the first 20 elements of a vector, or only the elements that are not NA, or only those that are positive or correspond to a specific variable of interest.

```{r, collapse=TRUE}
# Example object
c(1.45989479898373, NA, NA, NA, 0.343478985070138, -0.765250400381116, 
0.592711619887493, 0.475780864848148, NA, -1.49800584466182, 
0.970123875340488, NA, 1.49381622542397, NA, -1.12876916959833, 
NA, 0.901829750106227, NA, NA, -0.978058674564382, -0.14728923135403, 
NA, -0.0930325364474759, NA, NA, 0.265485649676846, NA, -0.681746435250382, 
0.368718594581899, NA, -1.43797529068594, -0.596616172600607, 
NA, NA, -1.16721680865855, NA, NA, NA, NA, 0.281394062459285)

# Get only the NAs
x[is.na(x)]

# Get all the non-NA values
y <- x[!is.na(x)]

# Get the 3rd, 5th and 7th element of x
x[c(3,5,7)]

# Get all elements of x EXCEPT the 2nd and 10th
x[c(-2, -10)] # Using negative integer indexes returns all the elements of x except the negatives specified
# A quicker way to do this is to put the negative sign in front of the vector
x[-c(2,10)]
```


# Subsetting lists

One of the reasons to use names for objects is to make reference to different elements of the object by name. 

```{r, collapse=TRUE}
x <- list(foo = 1:4, bar = 0.6)
# Extract the first element of the list x
# Use single brackets
x[1]

# Use dollar sign
x$foo

# Use double brackets
x[[1]]
```

foo is the first element and it is a sequence from 1 to 4, and bar is the second element and it is a number.
Using the single brackets we got the list that contained the sequence 1 to 4. By using the dollar sign and the double brackets we just got the sequence. That is their difference.

## Extracting multiple elements of a list
To extract multiple elements from a list we need to use the single bracket operator.

```{r, collapse=TRUE}
x <- list(foo = 1:4, bar = 0.6, baz = "hello")
# Extract the first and last element of the above list
x[c(1,3)]
```

You cannot use double brackets or the dollar sign operator to extract multiple elements.

# Subsetting matrices


```{r, collapse=TRUE}
x <- matrix(1:6, 2, 3)
x[1,2] #first row, second column

# indices can also be missing
x[1, ] # prints first row
x[, 2] # prints second column

```

When a single element of a matrix is retrieved, it is returned as a vector of length 1 rather than a 1 x 1 matrix. This behavior can be turned off by setting `drop = FALSE`.

```{r, collapse=TRUE}
x <- matrix(1:6, 2, 3)
x[1,2, drop = FALSE]
```
This gave us back a 1 x 1 matrix with the element 3 in it. An example with missing indices:

```{r, collapse=TRUE}
x <- matrix(1:6, 2, 3)
x[1, , drop = F]
```

# Subsetting with names - Partial Matching

A partial match occurs if the whole of the element of x matches the beginning of the element of table. Finally, all remaining elements of x are regarded as unmatched.

Partial matching of names is allowed with `[[]]` and `$`.

```{r, collapse=TRUE}
x <- list(aardvark = 1:5)
x$a
x[["a"]]
```
The `$` looks for a name in this list that matches the letter a. In this case we only have one element and (aardvark) and it returns the associated sequence (1:5).

The double bracket operator expects that the name you pass it is going to be an exact match for one of the names in the list. By default, the `[[]]` does not do partial matching. In the above list we get a NULL value because there is no element in the list that has the name a. However, if we pass a second argument (`exact = FALSE`) we can make the search more flexible:

```{r, collapse=TRUE}
x <- list(aardvark = 1:5)
x[["a", exact = FALSE]]
```
# Subsetting - Missing values (NAs)

### Removing missing values 

```{r, collapse=TRUE}
x <- c(1, 2, NA, 4, NA, 5)
bad <- is.na(x) #Creates logical vector specifying where the NAs are.
bad
x[!bad] #From x, it removes bad (which has the NAs).
```
Just like single-bracket sub-setting, saying `x[bad]` would return only the values in bad which are the NAs. Including the exclamation mark `x[!bad]` means that we want everything except what is stored in bad.

### Getting only the non-missing values

```{r, collapse=TRUE}
x <- c(1, 2, NA, 4, NA, 5)
good <- complete.cases(x)
good
x[good]
```
### Removing NAs from data frames

We can use complete.cases to remove missing values from data frames.

```{r, collapse=TRUE}
airquality[1:6, ] #data frame
good <- complete.cases(airquality)
airquality[good, ][1:6, ]
```
# Vectorized operations

Suppose that you have two vectors, x and y. x is the sequence 1:4 and y 6:9 and you want to add the two vectors together. What we mean when we say add them is to add the first element of x to the first element of y (i.e. 1 + 6). In other programming languages this would require to loop through each element but in R we can just use the + operator:

```{r, collapse=TRUE}
x <- 1:4; y <- 6:9
x + y
```
We can also multiply, subtract and divide.

```{r, collapse=TRUE}
x <- 1:4; y <- 6:9
x - y
x * y
x / y
```


##### Vectorized matrix operations

You can do the same with matrices.

```{r, collapse=TRUE}
x <- matrix(1:4, 2, 2); y <- matrix(rep(10, 4), 2, 2) # rep() replicates 10 four times

x * y ## element-wise multiplication
x / y 
x %*% y ## true matrix multiplication

```


