---
title: "Control Structures"
author: "Akis"
date: "24/01/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<center><h1> Control Structures </center></h1>

Control structures control the flow of execution of the program. Common structures are:

* if,else: testing a condition
* for: execute a loop a fixed number of times
* while: execute a loop *while* a condition is true
* repeat: execute an infinite loop
* break: break the execution of a loop
* next: skip an iteration of a loop
* return: exit a function

# if-else statements

Tets logic conditions...

```{r, eval=FALSE}
### If-Else ###
if(<condition>) {
  ## do something
} else {
  ## do something else
}

### Elif (in python) ###
if(<condition>) {
  ## do something
} else if(<condition2>) {
  ## do something different
} else {
  # do something different
}
```

The else part is optional but it can be used if you want to do something alternatively. Example:

```{r}
# example 1
x <- 2

if(x > 3) {
  y <- 10
} else {
  y <- 10
}

# example 2
y <- if(x > 3) {
  10
} else {
  0
}

# example 3
if(x > 1) {
  z <- 5
}
```



