---
title: "Control Structures"
author: "Akis"
date: "24/01/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<center><h1> Control Structures </center></h1>

Control structures control the flow of execution of the program. Common structures are:

* if,else: testing a condition
* for: execute a loop a fixed number of times
* while: execute a loop *while* a condition is true
* repeat: execute an infinite loop
* break: break the execution of a loop
* next: skip an iteration of a loop
* return: exit a function

<center><h1> if-else statements </center></h1>

Tets logic conditions...

```{r, eval=FALSE}
### If-Else ###
if(<condition>) {
  ## do something
} else {
  ## do something else
}

### Elif (in python) ###
if(<condition>) {
  ## do something
} else if(<condition2>) {
  ## do something different
} else {
  # do something different
}
```

The else part is optional but it can be used if you want to do something alternatively. Example:

```{r, collapse=TRUE}
# example 1
x <- 2

if(x > 3) {
  y <- 10
} else {
  y <- 10
}

# example 2
y <- if(x > 3) {
  10
} else {
  0
}

# example 3
if(x > 1) {
  z <- 5
}
```

<center><h1> For loops </center></h1>

For loops are most commonly used for iterating over the elements of an object.

```{r, collapse=TRUE}
for(i in 1:10) {
  print(i)
}
```

This loop takes the **i** variable and in each iteration of the loop gives it values 1,2,3, ..., 10, and then exits.

```{r, collapse=TRUE}
x <- c("a", "b", "c", "d")

# these two loops have the same behavior
for(i in 1:4) {
  print(x[i])
}

for(letter in x) {
  print(letter)
}
```

## Nested for loops

You can have a for loop inside a for loop; this is called nesting. Nessting beyond 2-3 levels canbe very difficult to understand.

```{r, collapse=TRUE}
x <- matrix(1:6, 2, 3)

for(i in seq_len(nrow(x))) {
  for(j in seq_len(ncol(x))) {
    print(x[i, j])
  }
}
```

<center><h1> While loops </center></h1>

While loops begin by testing a condition. If it is true, then they execute the loop body. once the loop body is executed, the condition is tested again, and so forth.

```{r, collapse=TRUE}
count <- 0
while(count < 10) {
  print(count)
  count <- count +1
}
```

### Conditions

Sometimes there will be more than one condition in the tet.

```{r, collapse=TRUE}
z <- 5

while(z >= 3 && z <= 10) { # && is and
  print(z)
  coin <- rbinom(1, 1, 0.5) # flip a coin, rbinom = random number generation
  
  if(coin == 1) { # random walk
    z <- z + 1
  } else {
    z <- z - 1
  }
}
```

<center><h1> Repeat, Next, Break </center></h1>

## repeat
Repeat initiates an infinite loop. The only way to exit a **repeat** loop is to call **break**.

```{r, eval=FALSE}
x0 <- 1
tol <- 1e-8

repeat {
  x1 <- computeEstimate() ## imaginary function
  
  if(abs(x1 - x0) < tol) {
    break
  } else {
    x0 <- x1
  }
}
```

## next, return

**next** is used to skip an iteration of a loop.

```{r, collapse=TRUE}
for(i in 1:100) {
  if(i <= 20) {
    ## Skip the first 20 iterations
    next
  }
  ## Do something here
}
```

return signals that a function should exit and return a given value.