---
title: "Dates and Times in R"
author: "Akis"
date: "02/02/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

R has developed a special representation of dates and times

* Dates are represented by the `Date` class
* Times are represented by the `POSIXct` or the `POSIXlt` class
* Dates are stored internally as the number of dates since 1970-01-01
* Times are stored internally as the number of seconds since 1970-01-01

<center> Dates </center>

Dates are represented by the **Date class** and can be coerced from a character string using the `as.Date()` function.

```{r}
x <- as.Date("1970-01-01")
x

unclass(x) 
unclass(as.Date("1970-01-02"))
```

Since x is the date 1970-01-01, unclass(x) returns zero because there have been zero days since 1970-01-01. unclass(as.Date("1970-01-02")) returns one because there has been one day since 1970-01-01. Any dates before 1970 would be represented as negative numbers.

<center> Times </center>

Times are represented using the POSIXct or the POSIXlt class.

* `POSIXct` is just a very large integer under the hood; it is a useful class when you want to store times in something like a data frame.
* `POSIXlt` is a list underneath and it stores a bunch of other useful information like the day of the week, day of the year, month, day of the month.

There are a number of generic functions that work on dates and times.

* **weekdays**: give the day of the week
* **months**: give the month name
* **quarters**: give the quarter number("Q1", "Q2", "Q3", or "Q4")

Times can be coerced from a character string using the **as.POSIXlt** or **as.POSIXct** function.

```{r}
x <- Sys.time() # system time
x

p <- as.POSIXlt(x) # remember POSIXlt is a list underneath
names(unclass(p))

p$sec # returns the seconds 
```

```{r, eval=FALSE}
x <- Sys.time()
x # Already in POSIXct format
unclass(x)

x$sec #Error: $ operator is invalid for atomic vectors
p <- as.POSIXlt(x)
p$sec
```

Finally, there is the `strptime` function in case your dates are written in a different format.

```{r}
datestring <- c("February 02, 2022 15:00", "December 10, 2011 9:00")
x <- strptime(datestring, "%B %d, %Y %H:%M")
x

class(x)
```

If you cannot remember the formatting strings check `?strptime` for details.

## Operations on Dates and Times

You can use mathematical operations on dates and times. Well, really just + and -. You can do comparisons too (i.e. ==, <=)

```{r, eval=FALSE}
x <- as.Date("2012-01-01")
y <- strptime("9 Jan 2011 11:34:21", "%d %b %Y %H:%M:%S")
x-y
## Warning: Incompatible methods ("-.Date", "-.POSIXt") for "-"
## Error in x - y : non-numeric argument to binary operator
x <- as.POSIXlt(x)
x-y
## Time difference of 356.3 days
```

