---
title: "Defining binary operators"
author: "Akis"
date: "03/02/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<center> Define your own binary operators </center>\

You're familiar with adding, subtracting, multiplying, and dividing numbers in R. To do this you use the +, -, *, and / symbols. These symbols are called binary operators because they take two inputs, an input from the left and an input from the right. In R you can define your own binary operators.

The syntax for creating new binary operators in R is unlike anything else in R, but it allows you to define a new syntax for your function. I would only recommend making your own binary operator if you plan on using it often!

User-defined binary operators have the following syntax:
     `%[whatever]%` 
where [whatever] represents any valid variable name.
 
Let's say I wanted to define a binary operator that multiplied two numbers and
then added one to the product. An implementation of that operator is below:

```{r,eval=F}

"%mult_add_one%" <- function(left, right){ # Notice the quotation marks!
  left * right + 1
}

```

I could then use this binary operator like `"4 %mult_add_one% 5"` which would
evaluate to 21.

Write your own binary operator below from absolute scratch! Your binary
operator must be called `%p%` so that the expression:

      "Good" %p% "job!"

will evaluate to: "Good job!"

```{r, collapse=TRUE}
"%p%" <- function(left, right){ 
  paste(left, right)
}

"Good" %p% "job"

"I" %p% "am" %p% "starting" %p% "to" %p% "like" %p% "this!"
```


