---
title: "Ellipsis"
author: "Akis"
date: "03/02/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The ellipses can be used to pass on arguments to other functions that are
used within the function you're writing. Usually a function that has the
ellipses as an argument has the ellipses as the last argument. The usage of
such a function would look like:

```{r, eval=FALSE}
ellipses_func(arg1, arg2 = TRUE, ...)
```

In the above example arg1 has no default value, so a value must be provided
for arg1. arg2 has a default value, and other arguments can come after arg2
depending on how they're defined in the ellipses_func() documentation.
Interestingly the usage for the paste function is as follows:

```{r, eval=FALSE}
paste (..., sep = " ", collapse = NULL)
```

Notice that the ellipses is the first argument, and all other arguments after
the ellipses have default values. This is a strict rule in R programming: all
arguments after an ellipses must have default values. Take a look at the
simon_says function below:

```{r, eval=FALSE}
simon_says <- function(...){
  paste("Simon says:", ...)
}
```

The simon_says function works just like the paste function, except the
begining of every string is prepended by the string "Simon says:"

Telegrams used to be peppered with the words START and STOP in order to
demarcate the beginning and end of sentences. Write a function below called 
telegram that formats sentences for telegrams.
For example the expression `telegram("Good", "morning")` should evaluate to:
"START Good morning STOP"

```{r, collapse=TRUE}
telegram <- function(...){
  paste("START", ... , "STOP")
}

telegram("Good", "morning", "Vietnam!")
```


<center> Unpack arguments from an ellipses </center>

Let's explore how to "unpack" arguments from an ellipses when you use the
ellipses as an argument in a function. Below I have an example function that
is supposed to add two explicitly named arguments called alpha and beta.

```{r,eval=FALSE}
add_alpha_and_beta <- function(...){
  # First we must capture the ellipsis inside of a list
  # and then assign the list to a variable. Let's name this
  # variable `args`.

  args <- list(...)

  # We're now going to assume that there are two named arguments within args
  # with the names `alpha` and `beta.` We can extract named arguments from
  # the args list by using the name of the argument and double brackets. The
  # `args` variable is just a regular list after all!
   
  alpha <- args[["alpha"]]
  beta  <- args[["beta"]]

  # Then we return the sum of alpha and beta.

  alpha + beta 
}
```

Have you ever played Mad Libs before? The function below will construct a
sentence from parts of speech that you provide as arguments. We'll write most
of the function, but you'll need to unpack the appropriate arguments from the
ellipses.

```{r, collapse=TRUE}
mad_libs <- function(...){
  # Do your argument unpacking here!
  my_args <- list(...)
  
  place <- my_args["place"]
  adjective <- my_args[["adjective"]]
  noun <- my_args[["noun"]]
  

  # Don't modify any code below this comment.
  # Notice the variables you'll need to create in order for the code below to
  # be functional!
  paste("News from", place, "today where", adjective, "students took to the streets in protest of the new", noun, "being installed on campus.")
}

mad_libs(place = "Cyprus", adjective = "hungry", noun = "gym")
```


